---
title: "Performant reverse geocoding"
author: "Hugh Parsonage"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Performant reverse geocoding}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r loadPackages}
library(scales)
library(ggplot2)
library(data.table)
library(hutils)
library(hutilscpp)
library(fst)
library(magrittr)
```

```{r loadPSMA}
library(PSMA)
```

```{r psma_env}
psma_env <- getOption("PSMA_env", new.env())
```

```{r ADDRESS_DETAIL_ID__by__LATLON}
dt <- "ADDRESS_DETAIL_ID__by__LATLON"
if (exists(dt, envir = psma_env)) {
  x <- get(dt, envir = psma_env, inherits = FALSE)
} else {
  x <- fst::read_fst(system.file("extdata", "address2.fst",
                                 package = "PSMA"),
                     as.data.table = TRUE)
  x[, "LATITUDE" := lat_int + lat_rem / 10^7]
  x[, "LONGITUDE" := lon_int + lon_rem / 10^7]
  assign(dt,
         value = x,
         envir = psma_env)
}
ADDRESS_DETAIL_ID__by__LATLON <- x
```

## Naively reverse geocoding is a bad idea

Consider the problem of geocoding the City of Melbourne's parking spaces.

```{r latlon_by_bayid}
latlon_by_bayid <- fread(system.file("extdata", "MEL-latlon_by_bayid.tsv", package = "PSMA"))
```

There are `r comma(nrow(ADDRESS_DETAIL_ID__by__LATLON))` addresses in the PSMA data base. With
`r nrow(latlon_by_bayid)` parking spaces to geocode, a naive cross join requires 
`r round(nrow(ADDRESS_DETAIL_ID__by__LATLON) * as.double(nrow(latlon_by_bayid)) / 1e9, 1)` billion
 distance calculations.

The package `hutils` provides a lightweight `haversine_distance` function that only uses
primitive functions. It's essentially as fast as it can be. Yet, if we extrapolate from the
cost of a naive way to reverse geocode the first parking bay,

```{r bench_naive_distance_cj}
bench_naive_distance_cj <- 
   bench::system_time({
     haversine_distance(latlon_by_bayid$lat[1], 
                        latlon_by_bayid$lon[1], 
                        ADDRESS_DETAIL_ID__by__LATLON$LATITUDE,
                        ADDRESS_DETAIL_ID__by__LATLON$LONGITUDE) %>%
       which.min
   })
bench_naive_distance_cj * nrow(latlon_by_bayid)
```

we see that this is not very satisfactory. 

## Trim and trim

We can do better with little work.

A priori, we know that these parking spots are in Melbourne. In particular,
we know that all the possible addresses are going to be in a small, compact 
area. So let's restrict the search to addresses within 6 minutes of a degree around 
the parking locations. 

```{r addresses_near_MEL}
bench::system_time({
  min_lat <- latlon_by_bayid[, min(lat)] - 0.1
  max_lat <- latlon_by_bayid[, max(lat)] + 0.1
  min_lon <- latlon_by_bayid[, min(lon)] - 0.1
  max_lon <- latlon_by_bayid[, max(lon)] + 0.1♠
  
  addresses_near_MEL <- 
    ADDRESS_DETAIL_ID__by__LATLON %>%
    .[LATITUDE %between% c(min_lat, max_lat)] %>%
    .[LONGITUDE %between% c(min_lon, max_lon)]
})
```

```{r bench_nearby_distance_cj}
bench_nearby_distance_cj <- 
   bench::system_time({
     haversine_distance(latlon_by_bayid$lat[1], 
                        latlon_by_bayid$lon[1], 
                        addresses_near_MEL$LATITUDE,
                        addresses_near_MEL$LONGITUDE) %>%
       which.min
   })
bench_nearby_distance_cj * nrow(latlon_by_bayid)
```

There is another potential for performance improvement that may be peculiar to the PSMA data:
Many of the addresses have identical latitude and longitude:

```{r uniqueN_latlon}
addresses_near_MEL %>%
  .[, lapply(.SD, round, 5)] %>%
  .[, lapply(.SD, uniqueN), .SDcols = -1] %>%
  .[, lapply(.SD, "/", nrow(addresses_near_MEL))] %>%
  .[, lapply(.SD, percent)] %>%
  kable
```



## Radix sorting?


## R* tree?


## Parallelize


## GPU?



Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
